---
title: "Day 7 Lecture"
output: html_document
---
# 0. Libraries
```{r}
library(ggplot2)
```

# 1. Review HW 4
### A. Combining values in columns colA and colB when they are "complementary" (i.e., one has NAs where the other has values)
```{r}
mydata <- data.frame(colA = c(NA, 1, 2, NA, NA), colB = c(45, NA, NA, 2, 99))
mydata

mydata$colC <- mydata$colA # create colC and put the values in colA into colC
mydata$colC[is.na(mydata$colC)] # not sure if we need this

# wherever column A is NA, put the value that's in column B (when the value in column A is NA) into column C:
mydata$colC[is.na(mydata$colA)] <- mydata$colB[is.na(mydata$colA)]
mydata
```

### B. Checking for and locating NAs
```{r}
exdata <- data.frame(colA = c(NA, 1, 2, NA, 4), colB = c(45, NA, NA, 2, 99), colC = c(66, 5, NA, 3, 5))
exdata
complete.cases(exdata)
exdata[!complete.cases(exdata),]
which(!complete.cases(exdata))
```

### C. Miscellaneous notes
- if you want to change the order of factors, use **factor()**, DON'T use reorder()!
- which() takes a logical and returns the locations where it's true

# 2. Review HW 5
### A. Transforming variables
```{r}
minpair <- read.csv("~/Documents/UMD and CASL/Classes/R Class/r_programming_bsos/scottdata/CognitionPaperFinalData.csv")
head(minpair)
ggplot(minpair, aes(WithinMinPairs)) + geom_histogram() + geom_vline(xintercept = 1000, color = "red")

ggplot(minpair, aes(log(WithinMinPairs))) + geom_histogram() 
```

### B. Changing graph colors
**See http://colorbrewer2.org/**  
- use scale_fill_brewer(palette = "Set#") to use colorbrewer for discrete variables 
- use scale_fill_manual(values = c("names")) to set values manually for discrete variables 
- other color info: http://research.stowers-institute.org/efg/R/Color/Chart/index.htm
- wes anderson color palette: https://github.com/karthik/wesanderson

Setting colors using color brewer
```{r}
ggplot(minpair, aes(log(WithinMinPairs))) + geom_histogram(aes(fill = Corpus)) + scale_fill_brewer(palette = "Set2")

ggplot(minpair, aes(log(WithinMinPairs))) + geom_histogram(aes(fill = Merged)) + scale_fill_brewer(palette = "Set2")

ggplot(minpair, aes(log(WithinMinPairs))) + geom_histogram(aes(fill = Corpus)) + scale_fill_brewer(palette = "PuOr")
```

Setting colors based on actual color names
```{r}
# ggplot(minpair, aes(log(WithinMinPairs))) + geom_histogram(aes(fill = Corpus)) + scale_fill_manual(values = c("firebrick", "steelblue"))
```

Setting colors based on RGB values
```{r}
ggplot(minpair, aes(log(WithinMinPairs))) + geom_histogram(aes(fill = Merged)) + scale_fill_manual(values = c("firebrick", rgb(27, 158, 119, maxColorValue = 255)))
```

Color charts!
```{r}
n=12
pie(rep(1,n), col=rainbow(n))
pie(rep(1,n), col=heat.colors(n))
pie(rep(1,n), col=terrain.colors(n))
pie(rep(1,n), col=topo.colors(n))
pie(rep(1,n), col=cm.colors(n))
barplot(seq(1:12), col=topo.colors(12))
?rainbow
?palette
```

RColorBrewer package
```{r}
library(RColorBrewer)

par(mar = c(0, 4, 0, 0))
display.brewer.all()
```

wesanderson package
- general info: https://github.com/karthik/wesanderson
- help file: http://cran.r-project.org/web/packages/wesanderson/wesanderson.pdf
```{r}
library(wesanderson)

namelist

ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) + 
  geom_point(size = 3) + 
  scale_color_manual(values = wes.palette(3, "Cavalcanti")) + 
  theme_minimal()
```

### B. Other visualization stuff
- plotting symbols: http://vis.supstat.com/2013/04/plotting-symbols-and-color-palettes/

# 3. Review Practice 6

# 4. Review HW 6

# 5. Other review

# 6. GitHub

# 7. GitHub Practice

# 8. Rest of the course
  - `dplyr`
  - functions
  - special topics
  - make-up time
