---
title: "Day 3 Lecture"
output: html_document
---
# Pulling from the repo
first time:
`git clone https://github.com/shoestringpsycholing/r_programming_bsos.git`

after that
`git pull origin master`


# Pulling from the repo

First time:
`git clone https://github.com/shoestringpsycholing/r_programming_bsos.git`

After that:
`git pull origin master`

# Some follow-up notes

## `<-` vs. `=`
<<<<<<< HEAD
both work!
not equivalent!

# Notebooks vs. scripts
Sweave
ctrl + shift + k knits
one hashtag is a header
two hashtags is a subheader
can collapse a section using the arrow next to a header/subheader

```{r}
x <- rnorm(10000)
hist(x)
```

The mean of 'x' is `r mean(x)`.

# Reading and writing data/files
Package for reading excel data:
  `xlsx`
  
Package for other stats packages (like SPSS, SAS, STATA):
  `foreign`
  
## "Reading" .RData
`load()` puts object in your workspace

```{r}

mydata <- laod("otherdata.RData") # otherdata.RData contains object "foo"

#mydata == "foo"
mydata <- foo # make a new object
```     

## Look at the options!
```{r}
?read.table

# mydata <- read.csv(filename, na.strings = c("NA", "N/A", "n/a", "", "nuthin"))
=======
Both work!
Not equivalent!

- the arrow (`<-`) is "more idiomatic" for R
- the single equals (`=`) is more common in other languages
- the arrow **always** means "assignment"
- the single equals does not mean assignment in some places (e.g., inside a function call)
- arguments for arrow:
  1. Visually distinctive
  2. Always unambiguously means "assignment"
- arguments for single equals:
  1. Less typing
  2. Common in many other programming languages
- *Final recommendation*: if you have a strong preference, just use that, if not, use arrow

# Notebooks vs. scripts
Sweave: the original "reproducible document", mix of LaTeX and R

R Markdown: one of the "cutting edge" formats, especially well-integrated into RStudio, that allows mixing of R code with text, document formatting, links, graphics, etc.

The following is a code chunk, and when `knitr()` is run on this document, it will display both the code and the output.

```{r}
x <- rnorm(10000)
hist(x)  # this is a histogram

t.test(x)
```

You can also insert "inline" results, like the following:

The mean of `x` is `r mean(x)`.

Clicking on the little question mark in the RStudio GUI above this document will point you towards lots of helpful documentation on the details of Markdown and R Markdown documents.

# Reading and writing data/files
## Weird file types
Package for reading Excel data:
  `xlsx`

Package for other stats packages (SPSS, SAS, Stata):
  `foreign`

## "Reading" .RData
`load()` puts objects in your workspace

```{r eval=FALSE}
# NOTE: this code block is marked eval = FALSE, so it will not
#       run. This is because the files mentioned are fake, and
#       will produce errors if run.
mydata <- load("otherdata.RData")  # otherdata.RData contains object "foo"

# mydata == "foo"
load("otherdata.RData")  # gets "foo" into your workspace
mydata <- foo # make a new object
```

## Look at the options!
```{r eval=FALSE}
?read.table

mydata <- read.csv(filename, na.strings = c("NA", "N/A", "n/a", " ", "nuthin"))
>>>>>>> e10bdf30634b899ceb31426b4d5af52accc940b6

?readLines
?scan
```

<<<<<<< HEAD
## Writing files!

```{r}
=======
## Writing files

```{r eval=FALSE}
>>>>>>> e10bdf30634b899ceb31426b4d5af52accc940b6
?write.table # .csv, etc.
```

Catches
<<<<<<< HEAD
 - rownames!
```{r}
write.table(mydata, file= "outputfile.txt", sep="/t", 
            row.names = FALSE)
=======
  - rownames!
```{r eval=FALSE}
write.table(mydata, file = "outputfile.txt", 
            sep = "\t",   # this makes the output tab-delimited
            row.names = FALSE)  # good for exporting data
>>>>>>> e10bdf30634b899ceb31426b4d5af52accc940b6
```

# Data types
   - numeric
   - character
   - factor
   - logical
   - NA
   - NULL
   - others (e.g, "Date")

## Numeric
<<<<<<< HEAD
  - Numbers!
  - sometimes integers can differ from non-integers
  - `as.numeric()` forces it to be a number (not integers, which are numbers with a special attribute)
  - '1L' is an integer
    - have means, medians
    
## Characters
  -`as.character()` converts to characters
  - more primitive than factors
  - can manipulate with regular expressions, other text functions

## Factors
  - Actually number under the hood
  - Act like numbers + labels
  - VERY useful for many knids of analyses
  - `stringsAsFactors` : option of how to read in character vectors
vectors
  - if a column is a *property* of data (not a measure itself), factors are usually best
  -- groupings = good as factors
  -- conditions = good as factors 
=======
  - numbers!
  - sometimes integers can differ from non-integers
  - `as.numeric()` forces it to be a number (not integers)
  - `1L` is an integer
  - have mean, medians

## Characters
  - `as.character()` converts to characters
  - more primitive than factors
  - can manipulate with regular expressions, other text functions
  
## Factors
  - Actually numbers under the hood
  - Act like numbers + labels
  - VERY useful for many kinds of analyses
  - if a column is a *property* of data (not a measure itself), factors are usually best
    - groupings = good as factors
    - conditions = good as factors
>>>>>>> e10bdf30634b899ceb31426b4d5af52accc940b6
  - if a column is *data*, maybe consider keeping it a string
  - `as.factor()` coerces to a factor
  - `factor()` creates a factor
  - factors have *levels*: categories
<<<<<<< HEAD
  - = FALSE if you don't want to read in characters as factors

## Logicals
  - TRUE and FALSE
  - if coerced to `as.numeric`, TRUE == 1, FALSE == 0
  -- can take the mean of a logical vector
  - Shortcuts: T and F
  -- *DANGER* : T and F can be overwritten
  - logical operators create logical vectors
  -- `==`
  -- `>`
  -- `<`
  -- `|`  inclusive Or
  -- `&`
  -- `>=`
  -- `!` not
  -- `&&` and `||` only look at the beginning ... for more specific programming situations
=======
  - `stringsAsFactors` : option for how to read in character vectors
    - = FALSE if you don't want to read in characters as factors
    
## Logicals
  - TRUE and FALSE
  - if coerced to `as.numeric`, TRUE == 1, FALSE == 0
    - can take the mean of a logical vector
  - Shortcuts: T and F
    - *DANGER*: T and F can be overwritten
  - logical operators create logical vectors
    - `==`
    - `>`
    - `<`
    - `|`  This is the "long bar", meaning (inclusive) "or" (aka "pipe")
    - `&`  == "and"
    - `>=`  greater or equal
    - `!`  is NOT
    - `&&` and `||`  for more specific programming sitauations
    
>>>>>>> e10bdf30634b899ceb31426b4d5af52accc940b6
```{r}
3 < 4

sleep
sleep$extra > 0

<<<<<<< HEAD
which(sleep$extra > 0) # says which of the things in the vector are true (ik ethe 6th one, the 8th one, etc.)
```

## NA and NULL
  - `NA` means "missing value", empty cell
  - `NULL` means "nothing", the absence of something
  - `Inf` and `-Inf` infinity and negative infinity
  - `NaN
=======
which(sleep$extra > 0)

```

## NA and NULL (and others)
  - `NA` mean "missing value" or "missing data"
  - `NULL` means "nothing", the absence of something
  - `Inf` and `-Inf`
  - `NaN`

Handling NAs
  - can be a a huge pain
  - sometimes `na.rm = TRUE` (like for `mean()`)
```{r}
x <- c(1, 2, NA, 4, 68)
mean(x)
mean(x, na.rm = TRUE)
```
  - sometimes it's other stuff
  - `cor()` use = "pairwise.complete.obs"
  - no universal "deal with NAs" argument
>>>>>>> e10bdf30634b899ceb31426b4d5af52accc940b6

# More subsetting
## Basics
```{r}
sleep
sleep[sleep$group == 1, ]
sleep$group == 1
sleep[which(sleep$group ==1), ]
which(sleep$group ==1)
sleep[sleep$group == 1 & sleep$extra > 0 , ]
sleep[sleep$group == 1 | sleep$extra > 0 , ]
sleep[sleep$group != 1, ]

sleep[sleep$ID == c(1, 3, 4, 6)] # DOESNT WORK
sleep[sleep$ID %in% c(1,3,4,6), ] # if in this set of things, returns TRUE, if not, returns FALSE
#checks if each number in set on left exists in vector on the right

?subset # check warning
subset(sleep, etra > 0 & grou p==2)
# dplyr is better (IMHO)
```
## Catches with NAs
```{r}
x <- c(1, 2, NA, 5)
x > 3

x == NA
is.na(x)
#?is.null use for programming 
```

## Basics
```{r}
sleep
sleep[sleep$group == 1, ]
sleep$group == 1
sleep[which(sleep$group == 1), ]
which(sleep$group == 1)
sleep[sleep$group == 1 & sleep$extra > 0, ]
sleep[sleep$group == 1 | sleep$extra > 0, ]
# "where"
sleep[!sleep$group == 1, ]

sleep[sleep$ID == c(1, 3, 4, 6), ]  # DOESN'T WORK
sleep[sleep$ID %in% c(1, 3, 4, 6), ]
sleep$ID %in% c(1, 3, 4, 6)
sleep[!sleep$ID %in% c(1, 3, 4, 6), ]  # "not in"

# ?subset  # check Warning
subset(sleep, extra > 0 & group == 2)
# dplyr is better (IMHO)

```

## Catches with NAs

```{r}
x <- c(1, 2, NA, 5)
x > 3

x == NA
is.na(x)
# ?is.null

```

# Data manipulation

## Combining data frame objects
<<<<<<< HEAD
  - rbind
=======
### `rbind()` and `cbind()`
- rbind
>>>>>>> e10bdf30634b899ceb31426b4d5af52accc940b6
    - glues rows on top of each other
    - expects same columns
    - new and improved `rbind_all()` and `rbind_list()` in `dplyr` package
  - cbind
    - glues columns next to each other
<<<<<<< HEAD
    - columns need to be the same length

  - merge
    - need something in common 
```{r}
subject.info <- data.frame(ID = 1:10, gender = c("male","female","male","male","female","female","female","female","male","female"))
sleep.merged <- merge(sleep, subject.info)
sleep
sleep.merged
nrow(sleep)
nrow(sleep.merged)

subject.info <- data.frame(ID = 1:12, gender = c("male","female","male","male","female","female","female","female","male","female", "male","male")) # what if you have two guys in the lab but no data for them
sleep.merged2 <- merge(sleep, subject.info, all=TRUE) # can't force factor levels on to the same factor

mysleep <- sleep
mysleep$ID <- as.numeric(mysleep$ID)
sleep.merged2 <- merge(mysleep, subject.info, all=TRUE)
sleep.merged2
```
    - simplest: have ONE column in common, named the same thing (with the samve values)
    - can get MUCH more complex
  - `nrow()` good check on what was gained/lost during a merge
  
## Creating new columns
```{r}
mysleep
mysleep$Study <- "Study1" # refer to a column that doesn't exist yet
mysleep$noise <- rnorm(20)
mysleep
```
## Transforming columns
=======
    - need to be the same length    
```{r}
sleep1 <- sleep[sleep$group == 1,  ]
sleep2 <- sleep[sleep$group == 2,  ]
sleep1
sleep2
sleep.reassembled <- rbind(sleep1, sleep2)
sleep.reassembled

sleep.left <- sleep[, 1:2]
sleep.left
sleep.ID <- sleep[, 3]
sleep.ID
sleep.reassembled.2 <- cbind(sleep.left, sleep.ID)
sleep.reassembled.2
```
>>>>>>> e10bdf30634b899ceb31426b4d5af52accc940b6

### `merge()`
```{r}
subject.info <- data.frame(ID = 3:12, gender = c("male", "male", "female", "female", "female", "female", "male", "female", "male", "male"))
summary(subject.info)

sleep.merged <- merge(sleep, subject.info)
sleep
sleep.merged
nrow(sleep)
nrow(sleep.merged)

mysleep <- sleep
mysleep$ID <- as.numeric(mysleep$ID)

sleep.merged2 <- merge(mysleep, subject.info, all = TRUE)
sleep.merged2

```
    - simplest: have ONE column in common, named the same thing (with the same values)
    - can get MUCH more complex
    - `nrow()` good check on what was gained/lost during a merge
    
## Creating new columns

```{r}
mysleep
mysleep$Study <- "Study1"
mysleep$noise <- rnorm(20)
mysleep
```

