---
title: "Day 4 Lecture"
output: html_document
---

# Review?

# Data manipulation
  1. Simple column transformations
  2. Dealing with factors
  3. Reshaping data
  4. Aggregation

## Simple column transformations
```{r}
mysleep <- sleep
mysleep$extra2 <- mysleep$extra * 2
mysleep$extra3 <- log(mysleep$extra + 10)
mysleep
mysleep$extra2 <- mysleep$extra2 - 4
mysleep
mysleep$extra2 <- NULL # deletes the column. can also subset.
mysleep
#inserted vector must be of the same length
mysleep$four <- c(1,2,3) #doesn't work
```

```{r}
mysleep <- sleep
mysleep$extra2 <- mysleep$extra * 2
mysleep$extra3 <- log(mysleep$extra + 10)
mysleep
mysleep$extra2 <- mysleep$extra2 - 4
mysleep
mysleep$extra2 <- NULL  # deletes the column
mysleep
mysleep$extra3 <- NA
mysleep
mysleep$three <- 3
mysleep
mysleep$four <- c(1, 2, 3) # doesn't work
mysleep
```

## Dealing with factors
  - levels
  - droplevels
  - recoding
  - reordering
  - as.numeric vs. as.character
  - stringsAsFactors = argument to `read.table()`

### droplevels
```{r}
mysleep <- sleep
summary(mysleep)
levels(mysleep$ID)
summary(mysleep$ID)

mysleep.1 <- subset(mysleep, group == "1")
mysleep.1 <- subset(mysleep.1, !ID == "3")
# mysleep.1 <- subset(mysleep, group == 1 & !ID == 3)
# mysleep.1 <- mysleep[mysleep$group == 1 & !mysleep$ID == 3, ]
summary(mysleep.1$ID)
summary(mysleep.1$group)

mysleep.2 <- droplevels(mysleep.1)
summary(mysleep.2$ID)
summary(mysleep.2$group)
```

### recoding & reordering
  - relevel, reorder, factor, levels
```{r}
mysleep <- sleep
summary(mysleep)

levels(mysleep$group)
levels(mysleep$group) <- c("pre", "post")  # changes LABELS of the levels
levels(mysleep$group)
mysleep

# levels(mysleep$group) <- c("post", "pre")  # DOESN'T CHANGE ORDER!
# levels(mysleep$group)
# mysleep

mysleep$group <- relevel(mysleep$group, "post")  # changes order
levels(mysleep$group)
summary(mysleep)
mysleep

# avoid reorder() except for special cases

mysleep$ID <- factor(mysleep$ID, levels = c("2", "3", "1", "5", "10", "7", "9", "8", "4", "6"))
levels(mysleep$ID)
mysleep

mysleep$IDlevelnum <- as.numeric(mysleep$ID)
mysleep

```

 - `reorder()`
  - takes two vectors, a factor, and an "ordering" vector
  - the "ordering" vector needs to be:
    - the same length
    - numeric values corrsponding to the order of levels in the factor

### droplevels
```{r}
mysleep <- sleep
levels(mysleep$ID)
summary(mysleep$ID)

mysleep.1 <- subset(mysleep, group== 1)
summary(mysleep.1) #even though there are no values in group that have 2, it will still remain as a level
mysleep.2 <- droplevels(subset(mysleep, group==1))
summary(mysleep.2)
mysleep.1 <- subset(mysleep.1, ID!= 3)
# mysleep.1 <- subset(mysleep, group == 1 & ID != 3) #alternate ways to do with subsetting
# mysleep.1 <- mysleep[mysleep$group == 1 & mysleep$ID != 3, ]
summary(mysleep.1$ID)
mysleep.2 <- droplevels(mysleep.1)
summary(mysleep.2$ID)
```

### recoding & reordering
  - relevel, reorder, factor, level
```{r}
mysleep <- sleep

#change your factors to NOT numbers

levels(mysleep$group)
levels(mysleep$group) <- c("pre","post") #changes LABELS of the levels
levels(mysleep$group)

mysleep$group <- relevel(mysleep$group, "post") #changes order
levels(mysleep$group)

# avoid reorder() except for special cases

mysleep$ID <- factor(mysleep$ID, levels = c("2","3","1","5","10","7","9","8","4","6")) # data stays the same, changes the factor level, good for regression... graphing... 
levels(mysleep$ID)
mysleep #data hasn't changed, same label associated with the same data
mysleep$IDlevelnum <- as.numeric(mysleep$ID) #force the factor to be a number, gives the factor number
mysleep
```

  - `reorder()`
  - takes two vectors, a factor, and an "ordering" vector
  - the "ordering" vector needs to be:
    - the same length
    - numeric values corresponding to the order of levels in the factor

  - as.numeric vs. as.character 
    - as.numeric will take factor numbers
    - use as.character

  - stringsAsFactors = argument to `read.table()'
    - if you don't want factors, read table in with stringsAsFactors = FALSE
    
  
## Reshaping data
<<<<<<< HEAD
Use the `reshape2` packages!!

```{r}
install.packages("reshape2")
=======
Use the `reshape2` package!!!

```{r}
# install.packages("reshape2")
>>>>>>> e3bad5ac3fb1f9c18fe80cdc46d0d06a2214b8c4
library(reshape2)
```

### "Long" vs. "wide" data
```{r}
mysleep <- sleep
<<<<<<< HEAD
levels(mysleep$group) <- c("pre","post")
sleep # is in long format already, measurement is all in one column (not across multiple columns)
mysleep.cast <- dcast(mysleep, ID ~ group, value.var = "extra") #left side of ~ is columns you want to stay columns, right side are columns youo want spread across multiple columns, value = data
mysleep.cast # now data from group 1 and then group 2 in separate columns
=======
mysleep
levels(mysleep$group) <- c("pre", "post")
mysleep.cast <- dcast(mysleep, ID ~ group, value.var = "extra")
mysleep.cast
>>>>>>> e3bad5ac3fb1f9c18fe80cdc46d0d06a2214b8c4

myff <- french_fries
head(myff)
?melt.data.frame
<<<<<<< HEAD
myff.melt <- melt(myff,id.vars = 1:4) #id.var are the columns you want to stay columns
myff.melt <- melt(myff, id.vars = c("time","treatment","subject","rep"))
myff.melt <- melt(myff, measure.vars = 5:9)
myff.melt <- melt(myff, measure.vars = c("potato","buttery","grassy","rancid","painty"))
head(myff.melt)

myff.melt <- melt(myff,id.vars = 1:4, value.name = "rating", variable.name = "taste")
head(myff.melt)

recastff <- dcast(myff.melt, time + treatment + subject + rep ~ taste, value.var = "rating")
head(recastff)
```
=======
myff.melt <- melt(myff, id.vars = 1:4)
myff.melt <- melt(myff, id.vars = c("time", "treatment", "subject", "rep"))
myff.melt <- melt(myff, measure.vars = 5:9)
myff.melt <- melt(myff, measure.vars = c("buttery", "potato", "grassy", "rancid", "painty"))
head(myff.melt)
summary(myff.melt)

myff.melt <- melt(myff, id.vars = 1:4, value.name = "rating", variable.name = "taste")
head(myff.melt)
summary(myff.melt)

recastff <- dcast(myff.melt, time + treatment + subject + rep ~ taste, 
                  value.var = "rating")
summary(recastff)
```

- Formula
>>>>>>> e3bad5ac3fb1f9c18fe80cdc46d0d06a2214b8c4

## Aggregation (with the `reshape2` package)
```{r}
aggregateff <- dcast(myff.melt, treatment ~ taste, value.var = "rating", fun.aggregate = mean, na.rm=TRUE)
aggregateff
```


```{r}
aggregateff <- dcast(myff.melt, treatment ~ taste, 
                  value.var = "rating", fun.aggregate = mean, na.rm = TRUE)

aggregateff
```

# Data cleaning

## A template for data cleaning
  1. Get the data to read in to R
  2. Fix column names
  3. Check all columns to make sure data is getting read in correctly
  4. Check NAs, decide what to do with them
  5. Remove excessive rows or other "defective" rows
  6. Re-code factors as needed
  
### Reading data
  - find a good `read.X()` function, where X = your type of data
    - `read.table()` is the basic one
    - `read.csv()` and `read.delim()` are convenience "wrappers" for `read.table()`, for CSV and tab-delimited files, respectively
    - `read.spss()` and `read.dta()` from the `foreign` package for SPSS and Stata files
  - for messier or more primitive kinds of files in weird formats, use `readLines()`, or `scan()` to pull in the data in a very raw format, which you will then need to format in R

### Fix column names

```{r}
mysleep <- sleep
oldnames <- colnames(mysleep)
<<<<<<< HEAD
mycolnames <- c("extrasleep","experimentalGroup","subjectID")
colnames(mysleep) <- mycolnames
data.frame(oldnames, mycolnames)
=======
mycolnames <- c("extraSleep", "experimentalGroup", "subjectID")
colnames(mysleep) <- mycolnames
data.frame(oldnames, mycolnames)

>>>>>>> e3bad5ac3fb1f9c18fe80cdc46d0d06a2214b8c4
```

### Check all columns
```{r}
```
### Check NAs
  - use summary()
<<<<<<< HEAD
  
=======

>>>>>>> e3bad5ac3fb1f9c18fe80cdc46d0d06a2214b8c4
### Remove rows/columns as needed

### Re-code factors as needed

# Complex objects
```{r}
myff <- french_fries
head(myff)
#want to do a regression predicting potato by treatment
results <- lm(potato ~ treatment, data = myff)
results
results.summary <- summary(results)
class(results)
plot(results)
#the results is a complex object
names(results.summary)
results.summary$coefficients
str(results)


```

```{r}
myff <- french_fries
results <- lm(potato ~ treatment, data = myff)
results
summary(results)
plot(results)
class(results)
names(results)
plot(results$residuals)
results.summary <- summary(results)
names(results.summary)
results.summary$coefficients
str(results)
class(results.summary$coefficients)

str(summary(results))
?lm
```


## Looking at the structure of objects

## Extracting bits and pieces

