---
title: "Day 6 Homework"
output: html_document
---
```{r}
mydata <- read.csv("Kowalski_withAgentProp.csv")
mydata <- mydata[ , c("Participant","Trial","Item","Condition","List","Correct","Time","agentprop")]
times <- read.csv("times.csv", row.names=NULL)
times$Condition <- as.factor(times$Condition)
levels(times$Condition) <- c("SRC","ORC","Act","Pass")
colnames(times) <- c("Order","Item","Condition","Region","Onset")
```
1. Think carefully about your data.  Think about something you would like to do multiple times, to subsets of your data.  

2. Write out (in plain English, in the notebook) what you are trying to do
I would like to code the data for the critical regions of the sentence stimuli. Each region and each sentence was a different length.

3. Write some "pseudo-code" comments in an R chunk, as an outline
```{r}
# if this list 
# and this trial (or item) 
# and greater than these times and less than those times,
# region x
```
4. Write some (real) code that will do what you want ONCE, on one subset of the data.
```{r}
my.vector <- vector(length = 288)

mydata.merged <- merge(times, mydata) 
summary(mydata.merged)

if(mydata.merged$List == 2
   & mydata$Item == 1
   & mydata$Time >= 619){
  my.vector <- mydata$
   }
``` 

5. Copy that code chunk, and convert it into a loop, which will do what you want over all the groups, concatenating the results.

```{r}
time.vector <- mydata$Time

my.vector <- vector(length = 288)
for(this.group in time.vector) {
  if (this.group >= times$time) {
    my.vector <- times$Region
  }
}
  
### this made R very angry. 

```

6. Copy that loop again, and at a condition (`if`) statement to add some further control about ignoring some groups or some other conditional operation.
