---
title: "Day 8 Homework"
output: html_document
---
```{r}
mydata <- read.csv("Kowalski_withAgentProp.csv")
mydata <- mydata[ , c("Participant","Trial","Item","Condition","List","Correct","Time","agentprop")]
times <- read.csv("times.csv", row.names=NULL)
times$Condition <- as.factor(times$Condition)
levels(times$Condition) <- c("SRC","ORC","Act","Pass")
head(times)

library("dplyr")
library("reshape2")
```
1. Try to recreate the results of Homework 6 (or your intended results) using the `dplyr` functions.
```{r}
mydata.merged <- merge(times, mydata)
summary(mydata.merged)

mydata.filtered <- filter(mydata.merged, Time > Onset & Time < End)

#This was a very long/strange way to go about it (maybe) but first I merged the datafile with the times file so that the List and Item would match up and there would be rows where a region was specified along with the onset and end of that region for that List and Item. Then I filtered out the rows where the actual Time was greater than the Onset but less than the end, so that the rows that were left had the correct region associated with them. That might not make much sense, but I think it worked.
```

2. Find another kind of analysis/transformation/summary stat that you want to apply across multiple groups in your data. Use the functions from `reshape2` and `dplyr` to get these results.

```{r}
head(mydata.filtered)
mydata.filtered %>%
  group_by(Condition, Region) %>%
  summarize(mean1 = mean(agentprop), SD1 = sd(agentprop)) 
```

3. For more practice, try another example of something like #2 (find another stat or grouping or something).
```{r}
mydata %>%
  group_by(Participant, Condition) %>%
  summarize(mean1 = mean(Correct)) %>%
  filter(mean1 > .70)
```
4. Do yet another case of #2 above, but this time throw in a transformation, too.
```{r}
mydata.filtered %>%
  group_by(Participant, Condition) %>%
  filter(Region == "Disamb") %>%
  mutate(mean.agentprop = mean(agentprop)) %>%

```