---
title: "Day 9 Homework"
output: html_document
---
```{r}
mydata <- read.csv("Kowalski_rprogrammingdata.csv")
library("dplyr")
```
1. Write a function to do a bunch of clean-up and formatting to your data.
```{r}
my.msec50.func <- function(data, time1, time2) {
  mutate(data, ms50 = floor((time1 - time2)/50))
}

mydata2 <- my.msec50.func(mydata, mydata$Time, mydata$StartTime)
head(mydata2)
```
2. Make sure this function works on your data.

3. Get some subset of your data, or find a similar-but-not-identical data set.  Change up the column order.  Re-run your function on this new data.  Does it work?
```{r}
mydata.rearrange <- mydata
mydata.rearrange$tempTime <- mydata$Time
mydata.rearrange$Time <- NULL

mydata3 <- my.msec50.func(mydata.rearrange, mydata.rearrange$tempTime, mydata.rearrange$StartTime)
head(mydata3)
````
4. Fix your function to be able to work on the new data.
it works!

5. Write a function to do some kind of summary stat, and use the `dplyr` functions to run this function over groups of your data.
```{r}
# summary stat function
agentprop.func <- function(x, y, z) { 
  output <- vector(length = length(x))  
  for (i in 1:length(x)) {
    if(x[i] == 1) {
      output[i] <- 1}
    if (y[i] == 1) {
      output[i]<- 0 }
    if (z[i] == 1) {
      output[i] <- 0.5}
    }
  mean.agentprop <- mean(output)
  mean.agentprop}

agentprop.func(mydata$Agent, mydata$Patient, mydata$Middle)

## with groups of my data

agentprop.func <- function(x, y, z) { 
  output <- vector(length = length(x))  
  for (i in 1:length(x)) {
    if(x[i] == 1) {
      output[i] <- 1}
    if (y[i] == 1) {
      output[i]<- 0 }
    if (z[i] == 1) {
      output[i] <- 0.5}
    }
  mean.agentprop <- mean(output)
  mean.agentprop}

mydata %>%
  filter(Condition == 1) %>%
  summarize(mean = agentprop.func(Agent, Patient, Middle))

mydata %>%
  group_by(Condition) %>%
  filter(Accuracy == 1) %>%
  summarize(mean = agentprop.func(Agent, Patient, Middle))

```