---
title: "Day 8 Practice"
output: html_document
---

1. Start with this data:

```{r}
library(dplyr)
library(reshape2)
minpair <- read.csv("scottdata/CognitionPaperFinalData.csv")
```

2. Get the means of `WithinMinPairs` for each level of `Corpus`
```{r}
head(minpair)
minpair.grouped <- group_by(minpair, Corpus)
head(minpair.grouped)
summarize(minpair.grouped, means = mean(WithinMinPairs))

## by pipe
minpair %>%
  group_by(Corpus) %>%
  summarize(mean = mean(WithinMinPairs))
```

3. Get cell means of `WithinMinPairs` by `Corpus` and `Merged`
```{r}
minpair %>%
  group_by(Corpus, Merged) %>%
  summarize(mean1 = mean(WithinMinPairs)) 
```

4. Melt down the last four columns, which are all frequency measures.  Get the min, max, and mean log values for all of these measures, by Corpus and Merged.
  - HINT: melt down the frequency columns first
```{r}
library("reshape2")
# mine:
# head(minpair)
# minpair.subset <- select(minpair, Corpus, Merged)
# minpair.melted <- melt(minpair, measure.vars = c("AllEntropy","WithinEntropy","SegEntropy","BiphoneEntropy"), value.name = "frequency", variable.name = "Entropy")
# head(minpair.melted)
# minpair.melted.2 <- melt(minpair.melted, measure.vars = c("HiSegTok","LoSegTok","HiSegType","LoSegType"), value.name = "frequency2", variable.name = "Seg")
# head(minpair.melted.2)
# minpair.melted.2 %>%
#   group_by(Corpus, Merged) %>%
#   summarize(min1 = min(frequency2), max1 = max(frequency2), mean.log = mean(log(frequency2)))
#####
minpair%>%
  melt(measure.vars = 14:17) %>%
  group_by(variable, Corpus, Merged) %>%
  summarize(minlog = min(log(value)), maxlog = max(log(value)), mean = mean(log(value)))

```
5. Do the same as 4, but only where HiSegTok is greater than the median value for each corpus.
```{r}
step1 <- minpair %>%
  group_by(Corpus) %>%
  summarize(medHST = median(HiSegTok))
step1

minpair %>%
  merge(step1) %>%
  filter(HiSegTok > medHST) %>%
  melt(measure.vars = 14:17) $>$
  group_by(variable, Corpus, Merged) %>%
  summarize(minlog = min(log(value)), maxlog = max(log(value)), mean = mean (log(value)))
  

```
